
---
title: "Tidy Modeling with R: Chapter 6 (Fitting Models with Parsnip)"
output:
  html_document: default
  pdf_document: default
---

```{r performance-setup, include = FALSE}
library(tidymodels)
library(tidyverse)
tidymodels_prefer()
```

This notebook works through the contents of chapter 6 of the book TMWR on fitting models with Parsnip.

The data used is the Ames house sale price dataset, which is available from the `tidymodels` package.
```{r data}
data(ames)
ames <- mutate(ames, Sale_Price = log10(Sale_Price))

set.seed(502)
ames_split <- initial_split(ames, prop = 0.80, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test  <-  testing(ames_split)
```

# Linear Regression with Parsnip

We construct a linear regression model object.
```{r}
ames_lm <-
    linear_reg() %>%
    set_engine("lm")
```

We can view the call to the underlying object using the `translate()` method.
```{r}
translate(ames_lm)
```

We can fit it using a formula via the `fit()` method.
```{r}
ames_form_fit <-
    ames_lm %>%
    fit(Sale_Price ~ Longitude + Latitude, data = ames_train)
tidy(ames_form_fit)
```

We can also fit the model using data via the `fit_xy()` method, where `x` is a tibble contain the independent variables (features) and `y` is a vector containing the dependent variable (response).
```{r}
ames_xy_fit <-
    ames_lm %>%
    fit_xy(
        x = ames_train %>% select(Longitude, Latitude),
        y = ames_train %>% pull(Sale_Price)
    )
tidy(ames_xy_fit)
```

The two methods produce exactly the same coefficients and associated statistics.
```{r}
all( tidy(ames_xy_fit) == tidy(ames_form_fit) )
```

# Random Forest Regression Example

We can construct a random forest regression model object using the same approach as we used for the linear regression, though we do need to call `set_mode()` to indicate that we're using the random forest model for regression rather than classification. 

We also specify two hyperparameters, the number of trees, and the number of data points required to make a split in a tree. Parsnip creates standard names for hyperparameters, so we would use the same names for `randomForest` as we do for `ranger` below.
```{r}
ames_rf <-
    rand_forest(trees = 1000, min_n = 5) %>%
    set_engine("ranger") %>%
    set_mode("regression")
```

We can view the call to the underlying object using the `translate()` method.
```{r}
translate(ames_rf)
```

# Extracting the Underlying Model

We can use `extract_fit_engine()` to obtain the fitted model object from the Parsnip object.
```{r}
ames_form_fit %>% extract_fit_engine()
```

This allows us to call methods on the underlying object like the linear model's `summary()` function.
```{r}
ames_form_fit %>% extract_fit_engine() %>% summary()
```

We can get just the coefficients too.
```{r}
ames_form_fit %>% extract_fit_engine() %>% summary() %>% coef()
```

However, the coefficients are stored in a matrix and the methods to get the fitted model parameters vary from model to model. To obtain model paramet `broom` package's `tidy()` method provides a standard way to get model parameters as a tibble from any type of model.
```{r}
tidy(ames_form_fit)
```


# Making Predictions with Parsnip

We use `predict()` function to use the fitted model to make predictions on new data. The output is a tibble with the predictions.
```{r}
ames_test_small <- ames_test %>% slice(1:5)
predict(ames_form_fit, new_data = ames_test_small)
```

As the order of rows is the same as the original dataset, it's easy to add columns with additional data, such as the actual sale price and 95% confidence intervals.

```{r}
ames_test_small %>% 
  select(Sale_Price) %>% 
  bind_cols(predict(ames_form_fit, ames_test_small)) %>% 
  bind_cols(predict(ames_form_fit, ames_test_small, type = "pred_int")) 
```

Predictions work the same way for all model types, as we can see with this regression tree model for the Ames housing dataset.
```{r}
tree_model <- 
  decision_tree(min_n = 2) %>% 
  set_engine("rpart") %>% 
  set_mode("regression")

tree_fit <- 
  tree_model %>% 
  fit(Sale_Price ~ Longitude + Latitude, data = ames_train)

ames_test_small %>% 
  select(Sale_Price) %>% 
  bind_cols(predict(tree_fit, ames_test_small))
```

# References

  1. Max Kuhn and Julia Silge, _Tidy Modeling with R_, chapter 6. https://www.tmwr.org/models.html (2022)
